---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

- Loading the dataset into R
```{r}
library(ggplot2)

if(!file.exists("activity.zip")) {
    url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
    download.file(url, "activity.zip")
}
data <- read.csv(unz("activity.zip", "activity.csv"), 
                 colClasses = c("integer", "Date", "integer"))
```


## What is mean total number of steps taken per day?

-  The total number of steps taken per day
```{r}
steps <- aggregate(steps~date, data, sum)
ggplot(steps, aes(steps)) + 
  geom_histogram(bins = 10, color = "white", fill = "#FF9999", alpha = 0.8) + 
  labs(title = "The total number of steps taken per day", x = "Steps")
```

```{r}
mean.steps <- mean(steps$steps)
median.steps <- median(steps$steps)
options(scipen=999)
```

- The __mean__ total number of steps taken per day: _`r mean.steps`_.
- The __median__ total number of steps taken per day: _`r median.steps`_.

## What is the average daily activity pattern?

```{r}
avg.steps <- aggregate(steps~interval, data, mean)
ggplot(avg.steps, aes(interval, steps)) + 
  geom_line(color = "#FF9999") +
  labs(title = "Average number of steps across all days", x = "5-minute interval identifiers", y = "Steps")
```

```{r}
max.interval <- avg.steps$interval[which.max(avg.steps$steps)]
```
- The 5-minute interval, that contains the __maximum__ average number of steps: _`r max.interval`_.


## Imputing missing values

```{r}
complete <- complete.cases(data)
missing.val <- sum(!complete)
```
- The total number of rows with NAs: _`r missing.val`_.

To filling all of the missing values in the dataset, I adopted the strategy: using mean of 5-minute interval

```{r}
na.index = which(is.na(data$steps))
na.interval = data[na.index, 3]
fill.steps = sapply(na.interval, function(x) { avg.steps[(avg.steps$interval==x), 2]})
data2 <- data
data2[na.index, 'steps'] <- fill.steps
steps2 <- aggregate(steps~date, data2, sum)

ggplot(steps2, aes(steps)) + 
  geom_histogram(bins = 10, color = "white", fill = "#FF9999", alpha = 0.8) + 
  labs(title = "The total number of steps taken per day \n (new dataset filled missing data)", x = "Steps")

mean2.steps <- mean(steps2$steps)
median2.steps <- median(steps2$steps)
```

###After filling the missing values

- The __mean__ total number of steps taken per day: _`r mean2.steps`_.
- The __median__ total number of steps taken per day: _`r median2.steps`_.

The mean number of steps taken per day are the same (_`r mean2.steps`_).
But, the median number of steps taken per day are slightly different (before filling missing data: _`r median.steps`_, after filling missing data: _`r median2.steps`_). It is probably due to filling missing values by the mean of 5 minute interval.


## Are there differences in activity patterns between weekdays and weekends?

```{r}
data2['day'] = factor(ifelse(as.POSIXlt(data2$date)$wday %in% c(0,6), "weekend", "weekday"))

avg2.steps <- aggregate(steps~interval + day, data2, mean)

ggplot(avg2.steps, aes(interval, steps)) + 
  geom_line(color = "#FF9999") +
  labs(title = "Average steps by day", x = "5-minute interval identifiers", y = "Steps") +
  facet_grid(day ~ .)
```
